---
title: "Malnutritional analysis"
output: github_document
---

This model uses code developed by Nicky Welton to estimate the assocaition between malnutrition and pneumonia mortality. 

```{r setup}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```


Packages 

```{r}
library(tidyverse)
library(rjags)
```


# Run model with example data

First check the model runs succesfully in Jags using example data.

## Example data

Need to convert the example data into matrices and vectors for JAGS.

```{r example_data}
metadata <- list(ng=3,ns=24)
ex_df <- read_delim("Supporting/example_data.txt", delim = "\t")
matrices <- list(r = NA, n = NA, g = NA)
matrices[] <- map(names(matrices), ~ ex_df %>% 
                  select(starts_with(.x), -na) %>% 
                  as.matrix())
na <- ex_df %>% pull(na)
pi2 <- ex_df %>% pull(pi2)


list_data <- c(matrices, 
               list(na = na, pi2 = pi2),
               metadata)
knitr::kable(ex_df)
```


## Run model

The following is the BUGS/JAGS code for the fixed effects model.

```{r getcode}
a <- read_lines("Supporting/FE_model.txt")
cat(a)
```

## Initialise model and return results

Very small number of iterations so can run code quickly for now. Add more later.

```{r runexamplemodel}
mod_example <- jags.model(file = "Supporting/FE_model.txt",
                            data = list_data, n.chains = 1, n.adapt = 1000)
update(mod_example, 100)

data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod_example, c("d"),n.iter = 100)
summary(LINE.out)
```

## Process real data into format for analysis

Process real data, in the first instance assume that the proportion in each category is known for all data.

First rename variables and recode. There should be a maximum of three groups for each study/malnutrition category type combination.

## Categorise pattern of collapsing for each study/manutrition category combination

Calculate which studies have missing event data, and which have missing "n" data.

```{r pat_maln}
mort <- read_csv("Data/Mortality_Numbers.csv")
names(mort) <- str_to_lower(names(mort)) 
names(mort) <- str_replace_all(names(mort), " ", "_")

maln_cat3 <- c("None", "Moderate", "Severe")

all_present <- mort %>% 
  group_by(study, malnutrition_category) %>% 
  summarise(all_n = all(maln_cat3 %in% malnutrition_severity) & !any(is.na(total)),
            all_r = all(maln_cat3 %in% malnutrition_severity) & !any(is.na(died))) %>% 
  ungroup()

unique(mort$malnutrition_severity) %>% dput()

```


```{r maln_data}
mort_slct <- mort %>% 
  select(study, malnutrition_category, n = total, r = died, maln = malnutrition_severity)

grp_lbls <-  mort_slct %>% 
  group_by(study, malnutrition_category) %>% 
  mutate(g = seq_along(study)) %>% 
  ungroup() %>% 
  mutate(g_lbl = case_when(
    maln == "None" ~ 1L,
    maln == "Moderate" ~ 2L,
    maln == "Severe" ~ 3L,
    maln == "Non-Severe" ~ 12L,
    maln == "All" ~ 23L)
  )

```

## Calculate the proportions 

Calculate the proportion in the second category within each collapsed category. Having done so collapse the Ns. Where the proportion is unknown, assume it is the same as the mean.

First need to classify which are collapsed.

```{r calculatenproprs}


pi2 <- grp_lbls %>% 
  select(study, malnutrition_category, g_lbl, n) %>% 
  mutate(g_lbl = paste0("g", g_lbl),
         n = as.integer(n)) %>% 
  spread(g_lbl, n)

g2_1 <- ns %>% 
  filter(!is.na(g1), !is.na(g2)) %>% 
  summarise(res = mean (g2 / (g1 + g2))) 

g2_3 <- ns %>% 
  filter(!is.na(g2), !is.na(g3)) %>% 
  summarise(res = mean (g2 / (g2 + g3))) 


pi2 <- pi2 %>% 
  mutate(pi2 = case_when(
    is.na(g3) ~ 
  ))


```

Take the first malnutrition category for each study. Will need to ask group to make a decision on which to use. Then spread the dataframe to wide, so that we have a matrix of N's, events and group labels as per the earlier structure.


```{r labe_groups}
grp_lbls2 <- grp_lbls %>% 
  select(study, malnutrition_category, r, n, g, g_lbl) %>% 
  distinct(study, g, .keep_all = TRUE)

grp_lbls2n <- grp_lbls2 %>% 
  select(-r, -g_lbl) %>% 
  spread(g, n)

grp_lbls2n <- grp_lbls2 %>% 
  select(-r, -g_lbl) %>% 
  spread(g, n)

grp_lbls2r <- grp_lbls2 %>% 
  select(-n, -g_lbl) %>% 
  spread(g, r)

grp_lbls2g <- grp_lbls2 %>% 
  select(-n, -r) %>% 
  spread(g, g_lbl)

```

Check that the restructuring has kept the order of the studies.

```{r check_ordering}

identical(grp_lbls2n %>% select(1:2),
          grp_lbls2g %>%  select(1:2))

identical(grp_lbls2n %>% select(1:2),
          grp_lbls2r %>%  select(1:2))

```



