---
title: "Malnutrition analysis"
output: github_document
params:
  restrict: "all"
---

This model uses code developed by Nicky Welton to estimate the association between malnutrition and pneumonia mortality. 

```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```


Packages 

```{r}
library(tidyverse)
library(rjags)
library(mcmcplots)
library(ggplot2)
library(R2OpenBUGS)
```

# Nomenclature

For all analyses the malnutrition categories are interpreted as folows:-

coding|label
------|-------
1|None
2|Moderate
3|Severe
12|None or moderate
23|Moderate or severe

# Fixed effects model

## Run model with example data

First check the model runs successfully in Jags using example data.

Need to convert the example data into matrices and vectors for JAGS.

```{r example_data}
metadata <- list(ng=3,ns=24)
ex_df <- read_delim("Supporting/example_data.txt", delim = "\t")
matrices <- list(r = NA, n = NA, g = NA)
matrices[] <- map(names(matrices), ~ ex_df %>% 
                  select(starts_with(.x), -na) %>% 
                  as.matrix())
na <- ex_df %>% pull(na)
pi2 <- ex_df %>% pull(pi2)


list_data <- c(matrices, 
               list(na = na, pi2 = pi2),
               metadata)
knitr::kable(ex_df)
```


## Run model

The following is the BUGS/JAGS code for the fixed effects model.

```{r getcode}
a <- read_lines("Supporting/FE_model.txt")
print(a)
```

## Check model runs with example data

```{r runexamplemodel}
mod_example <- jags.model(file = "Supporting/FE_model.txt",
                            data = list_data, n.chains = 1, n.adapt = 1000)
update(mod_example, 100)

data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod_example, c("d"),n.iter = 100)
summary(LINE.out)
```

## Process real data into format for analysis

Process real data, in the first instance assume that the proportion in each category is known for all data.

First rename variables and recode. There should be a maximum of three groups for each study/malnutrition category type combination.

## Categorise pattern of collapsing for each study/manutrition category combination

Calculate which studies have missing event data, and which have missing "n" data. In the present analysis, any with complete n (ie count) data have complete events data.


```{r pat_maln}
mort <- read_csv("Data/Mortality_Numbers.csv")
names(mort) <- str_to_lower(names(mort)) 
names(mort) <- str_replace_all(names(mort), "\\s|\\-", "_")

if(params$restrict == "pre"){
  mort <- mort %>% 
    filter(pre_or_post_2000 == "pre")
}

if(params$restrict == "post"){
  mort <- mort %>% 
    filter(pre_or_post_2000 == "post")
}

mort <- mort %>% 
  select(-pre_or_post_2000)

mort <- mort %>% 
  mutate_at(vars(survived, died, total), as.integer)

mort <- mort %>% 
  group_by(study, measure) %>% 
  mutate(all_missing = all(is.na(total))) %>% 
  ungroup() %>% 
  filter(!all_missing) %>% 
  select(-all_missing)

maln_cat3 <- c("None", "Moderate", "Severe")

mort <- mort %>% 
  group_by(study, malnutrition_category) %>% 
  mutate(all_n =  all(maln_cat3 %in% malnutrition_severity) & !any(is.na(total)),
            all_r =  all(maln_cat3 %in% malnutrition_severity) & !any(is.na(died)),
            col12 = !all(maln_cat3 %in% malnutrition_severity) & any(malnutrition_severity == "All"),
            col23 = !all(maln_cat3 %in% malnutrition_severity) & any(malnutrition_severity == "Non-Severe")) %>% 
  ungroup() 

all_present_smry <- mort %>% 
  distinct(study, malnutrition_category, .keep_all = TRUE) %>% 
  group_by(all_n, all_r, col12, col23) %>% 
  summarise(studies = sum(!duplicated(study)),
            studies_times_categories = n())
col12 <- all_present_smry$studies[all_present_smry$col12]
col23 <- all_present_smry$studies[all_present_smry$col23]
knitr::kable(all_present_smry)
```


Of those studies with collapsed data, `r col12` collapse into none/moderate and `r col23` collapse into moderate severe.

## Calculate the proportion in group 2 for the collapsed studies

Since all of those with missing event data have missing totals for each category, we can only examine the proportion in group 2 where there is complete data. Calculate the proportion in the second category (moderate malnutrition) within each collapsed category. Having done so collapse the Ns. Where the proportion is unknown, assume it is the same as the mean proportion.

First need to classify which are collapsed.

```{r calculatenproprs}
pi2 <- mort %>% 
  filter(all_n) %>% 
  select(study, malnutrition_category, malnutrition_severity, total) %>% 
  mutate(total = as.integer(total)) %>% 
  spread(malnutrition_severity, total) %>% 
  mutate(pi2_coll12 = Moderate/(Moderate + None),
         pi2_coll23 = Moderate/(Moderate + Severe)) %>% 
  select(study, malnutrition_category, pi2_coll12, pi2_coll23)


pi2_for_mdl <- mort %>% 
  filter(all_n) %>% 
  select(study, malnutrition_category, malnutrition_severity, total) %>% 
  mutate(total = as.integer(total)) %>% 
  spread(malnutrition_severity, total) %>% 
  group_by(study, malnutrition_category) %>% 
  summarise_all(sum) %>% 
  ungroup()

cat_names <- unique(pi2_for_mdl$malnutrition_category) %>%  sort()
pi2_for_mdl <- map(cat_names, ~ pi2_for_mdl %>% 
                     filter(malnutrition_category == .x) %>% 
                     select(None, Moderate, Severe) %>% 
                     as.matrix())
names(pi2_for_mdl) <- cat_names

pi2_smry <- pi2 %>% 
  group_by(malnutrition_category) %>% 
  summarise_at(vars(pi2_coll12, pi2_coll23), mean) %>% 
  ungroup()

pi2_lng <- pi2 %>% 
  gather("collapsed_categories", "value", pi2_coll12, pi2_coll23) 
plot_dist <- ggplot(pi2_lng, aes(x = value, fill = collapsed_categories)) + geom_histogram() +
  facet_wrap(collapsed_categories ~ malnutrition_category)
plot_dist
```

The plot above shows the proportion of participants category 2 of 1 + 2 and in category 2 of 2+3 for all three types of measure. The proportion in 2 (moderate) is sometimes surprisingly low or high. 

```{r maln_data}
mort_slct <- mort %>% 
  select(study, malnutrition_category, n = total, r = died, maln = malnutrition_severity,
         col12, col23) %>% 
  inner_join(pi2_smry) %>%
  mutate(pi2 = case_when(
    col12 ~ pi2_coll12,
    col23 ~ pi2_coll23,
    TRUE ~ 1
  ))
  

grp_lbls <-  mort_slct %>% 
  group_by(study, malnutrition_category) %>% 
  mutate(g = seq_along(study)) %>% 
  ungroup() %>% 
  mutate(g_lbl = case_when(
    maln == "None" ~ 1L,
    maln == "Moderate" ~ 2L,
    maln == "Severe" ~ 3L,
    maln == "Non-Severe" ~ 12L,
    maln == "All" ~ 23L)
  )

```


First run for the "w/a" definition of malnutrition.

This code rearranges the dataframe  so that we have a matrix of N's, events and group labels as per the earlier structure.


```{r labe_groups}
grp_lbls2 <- grp_lbls %>% 
  select(study, malnutrition_category, r, n, g, g_lbl, pi2) %>% 
  filter(malnutrition_category == "w/a") %>% 
  distinct(study, g, .keep_all = TRUE)

grp_lbls2n <- grp_lbls2 %>% 
  select(-r, -g_lbl) %>% 
  spread(g, n)

grp_lbls2n <- grp_lbls2 %>% 
  select(-r, -g_lbl) %>% 
  spread(g, n)

grp_lbls2r <- grp_lbls2 %>% 
  select(-n, -g_lbl) %>% 
  spread(g, r)

grp_lbls2g <- grp_lbls2 %>% 
  select(-n, -r) %>% 
  spread(g, g_lbl)

grp_lbls2_col12 <- grp_lbls2 %>% 
  group_by(study) %>% 
  summarise(res = any(12 %in% g_lbl)) %>% 
  pull(res)

```

Check that the restructuring has kept the order of the studies. The following tests should be TRUE.

```{r check_ordering}

identical(grp_lbls2n %>% select(1:2),
          grp_lbls2g %>%  select(1:2))

identical(grp_lbls2n %>% select(1:2),
          grp_lbls2r %>%  select(1:2))

```


## Try running models on real data

Simplest model, assume that the proportion with moderate malnutrition of those with none-moderate, and the proportion with moderate of those with moderate-severe in those studies with collapsed data is the same as the average for those studies where there data are not collapsed.

```{r trymodel}
na <- 3 - apply(grp_lbls2n %>% select(`1`, `2`, `3`) %>% as.matrix(), 1, function(x) x %>% 
        as.integer() %>% 
        is.na() %>% 
        sum())
list_data2 <- list(r = grp_lbls2r %>% select(`1`, `2`, `3`) %>% as.matrix(),
                   n = grp_lbls2n %>% select(`1`, `2`, `3`) %>% as.matrix(),
                   g = grp_lbls2g %>% select(`1`, `2`, `3`) %>% as.matrix(),
                   na = na,
                   pi2 = grp_lbls2$pi2,
                   ng = 3,
                   ns = nrow(grp_lbls2n))

```


```{r runmodel}
mod1 <- jags.model(file = "Supporting/FE_model.txt",
                            data = list_data2, n.chains = 2, n.adapt = 1000)
update(mod1, 1000)
data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod1, c("d"),n.iter = 1000)
summary(LINE.out)
a <- summary(LINE.out)
a <- a$statistics[,"Mean"]
a <- round(a,2)
```

The above tables show that the log odds ratio is `r a[2]` for moderate versus none and `r a[3]` for severe versus none.

## Next step

Next step will be to estimate the proportion of people in category two rather than assume it is fixed. Can estimate this from the data or can get input from subject-matter experts about the likely proportion. The following estimates it from the data.

In order to prepare for this final decision we developed two models. One model assumes (approach one) that the proportion of none-moderate which are moderate lies, with equal probability, somewhere between 0 and 1 and that the proportion of moderate-severe which are moderate lies between 0 and 1 (ie a uniform prior). A second model (approach two) assumes that the proportion in each category is exchangeable between studies, using a random effects model to estimate the proportion for those studies where it is not recorded.

These models (which for speed only run for a small number of samples) both give similar result.

### Approach One

Add an indicator variables for whether we want proportion in category two of 1 and 2, or of 2 and 3. Where there is no missing data, the proportion of two in 2/3 will be calculated, but this is not used in the code.

In the first instance explore the use of a flat distribution for the proportion in category two. This is run in both JAGS and OPENBUGS, giving the same results.

```{r prop_model}
list_data3 <- list_data2
list_data3[["pi2"]] <- NULL
list_data3[["coll12"]] <- grp_lbls2_col12
mod2 <- jags.model(file = "Supporting/FE_model_estimate_prop.txt",
                            data = list_data3, n.chains = 1, n.adapt = 1000)
update(mod2, 1000)

data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod2, c("d", "theta1", "theta2"),n.iter = 1000)
summary(LINE.out)

inits <- function() {
  list(d = c(NA,-2,1),
       mu = rep(0, nrow(list_data3$r)))
}
list_data3_openbugs <- list_data3
list_data3_openbugs$coll12 <- as.integer(list_data3_openbugs$coll12)
mode2_bugs <- bugs(data = list_data3_openbugs, 
                   parameters.to.save = c("d", "theta1", "theta2"), n.iter = 1000,
                   inits = inits,
                   model.file = "FE_model_estimate_prop_openbugs.txt")
print(mode2_bugs)
```

Show get similar result with Dirichlet prior, just a check on Dirichlet coding. **Not sure that I (DM) have coded this correctly.**

```{r prop_model_dirich}
list_data3 <- list_data2
list_data3[["pi2"]] <- NULL
list_data3[["coll12"]] <- grp_lbls2_col12
mod2 <- jags.model(file = "Supporting/FE_model_estimate_prop_dirich.txt",
                            data = list_data3, n.chains = 1, n.adapt = 1000)
update(mod2, 1000)

data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod2, c("d", "theta"),n.iter = 1000)
summary(LINE.out)

```


### Approach Two

This is the modelling where we estimate the proportion in the moderate category using the data on proportions from other studies.


```{r prop_model_estimate}
pi2_choose <- pi2_for_mdl$`w/a`
pi2_choose12 <- rowSums(pi2_choose[, 1:2])
pi2_choose23 <- rowSums(pi2_choose[, 2:3])

list_data4 <- list_data3
list_data4[["n_complete2a"]]   <- pi2_choose[,2]
list_data4[["n_complete2b"]]   <- pi2_choose[,2]
list_data4[["n_complete12"]] <- pi2_choose12
list_data4[["n_complete23"]] <- pi2_choose23
list_data4[["ns_complete"]] <- length(pi2_choose12)


mod3 <- jags.model(file = "Supporting/FE_model_estimate_prop2.txt",
                            data = list_data4, n.chains = 2, n.adapt = 1000)
update(mod3, 1000)
# mcmcplot(as.mcmc(mod3))
data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod3, c("d", "theta1_prop", "theta2_prop"),n.iter = 1000)
summary(LINE.out)
```



# Random effects model - RE for effects but not for proportions

Return to approach 1, but re-run the model assuming random effects model for the effect of malnutrition on mortality.

```{r re_model}
mode3_bugs <- bugs(data = list_data3_openbugs, 
                   parameters.to.save = c("d", "theta1", "theta2"), n.iter = 10000,
                   inits = inits,
                   model.file = "RE_model_estimate_prop_openbugs.txt")
print(mode3_bugs)
```


# RE for effects and for proportions

Combine approaches 1 and 3, where assuming random effects for effect of malnutrition, and random effects for proportion in each malnutrition category.

Uncomment  `mcmcplot(as.mcmc(mode4_bugs))` to see model diagnostics.Look reasonable to me without too much autocorrelation. Chains appear to converge.

```{r re_model_re_prop}
list_data4_openbugs <- c(list_data3_openbugs, list_data4[c("n_complete2a", "n_complete2b", 
                                                         "n_complete12", "n_complete23", 
                                                         "ns_complete")])
mode4_bugs <- bugs(data = list_data4_openbugs, 
                   parameters.to.save = c("d", "theta1_prop", "theta2_prop"), n.iter = 10000,
                   inits = inits,
                   model.file = "RE_model_RE_estimate_prop_openbugs.txt")
# mcmcplot(as.mcmc(mode4_bugs))
print(mode4_bugs)
```
