---
title: "Malnutritional analysis"
output: github_document
---

This model uses code developed by Nicky Welton to estimate the association between malnutrition and pneumonia mortality. 

```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```


Packages 

```{r}
library(tidyverse)
library(rjags)
library(ggplot2)
```

# Fixed effects model

## Run model with example data

First check the model runs succesfully in Jags using example data.

Need to convert the example data into matrices and vectors for JAGS.

```{r example_data}
metadata <- list(ng=3,ns=24)
ex_df <- read_delim("Supporting/example_data.txt", delim = "\t")
matrices <- list(r = NA, n = NA, g = NA)
matrices[] <- map(names(matrices), ~ ex_df %>% 
                  select(starts_with(.x), -na) %>% 
                  as.matrix())
na <- ex_df %>% pull(na)
pi2 <- ex_df %>% pull(pi2)


list_data <- c(matrices, 
               list(na = na, pi2 = pi2),
               metadata)
knitr::kable(ex_df)
```


## Run model

The following is the BUGS/JAGS code for the fixed effects model.

```{r getcode}
a <- read_lines("Supporting/FE_model.txt")
print(a)
```

## Check model runs with example data in JAGS

Very small number of iterations so can run code quickly for now. Add more later.

```{r runexamplemodel}
mod_example <- jags.model(file = "Supporting/FE_model.txt",
                            data = list_data, n.chains = 1, n.adapt = 1000)
update(mod_example, 100)

data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod_example, c("d"),n.iter = 100)
summary(LINE.out)
```

## Process real data into format for analysis

Process real data, in the first instance assume that the proportion in each category is known for all data.

First rename variables and recode. There should be a maximum of three groups for each study/malnutrition category type combination.

## Categorise pattern of collapsing for each study/manutrition category combination

Calculate which studies have missing event data, and which have missing "n" data. IN the present analysis, any with complete n data have complete events data.

```{r pat_maln}
mort <- read_csv("Data/Mortality_Numbers.csv")
names(mort) <- str_to_lower(names(mort)) 
names(mort) <- str_replace_all(names(mort), " ", "_")

mort <- mort %>% 
  mutate_at(vars(survived, died, total), as.integer)

mort <- mort %>% 
  group_by(study, measure) %>% 
  mutate(all_missing = all(is.na(total))) %>% 
  ungroup() %>% 
  filter(!all_missing) %>% 
  select(-all_missing)

maln_cat3 <- c("None", "Moderate", "Severe")

mort <- mort %>% 
  group_by(study, malnutrition_category) %>% 
  mutate(all_n =  all(maln_cat3 %in% malnutrition_severity) & !any(is.na(total)),
            all_r =  all(maln_cat3 %in% malnutrition_severity) & !any(is.na(died)),
            col12 = !all(maln_cat3 %in% malnutrition_severity) & any(malnutrition_severity == "All"),
            col23 = !all(maln_cat3 %in% malnutrition_severity) & any(malnutrition_severity == "Non-Severe")) %>% 
  ungroup() 

all_present_smry <- mort %>% 
  distinct(study, malnutrition_category, .keep_all = TRUE) %>% 
  group_by(all_n, all_r, col12, col23) %>% 
  summarise(studies = sum(!duplicated(study)),
            studies_times_categories = n())
col12 <- all_present_smry$studies[all_present_smry$col12]
col23 <- all_present_smry$studies[all_present_smry$col23]
knitr::kable(all_present_smry)
```


Of those studies with collapsed data, `r col12` collapse into none/moderate and `r col23` collapse into moderate severe.

## Calculate the proportion in group 2 for the collapsed studies

Since, in the sample so far, all of those with missing event data have missing totals for each category, we need to examine the proportions in group 2 where these 

Calculate the proportion in the second category within each collapsed category. Having done so collapse the Ns. Where the proportion is unknown, assume it is the same as the mean.

First need to classify which are collapsed.

```{r calculatenproprs}
pi2 <- mort %>% 
  filter(all_n) %>% 
  select(study, malnutrition_category, malnutrition_severity, total) %>% 
  mutate(total = as.integer(total)) %>% 
  spread(malnutrition_severity, total) %>% 
  mutate(pi2_coll12 = Moderate/(Moderate + None),
         pi2_coll23 = Moderate/(Moderate + Severe)) %>% 
  select(study, malnutrition_category, pi2_coll12, pi2_coll23)


pi2_for_mdl <- mort %>% 
  filter(all_n) %>% 
  select(study, malnutrition_category, malnutrition_severity, total) %>% 
  mutate(total = as.integer(total)) %>% 
  spread(malnutrition_severity, total) %>% 
  group_by(study, malnutrition_category) %>% 
  summarise_at(vars(-study, -malnutrition_category), sum) %>% 
  ungroup()

cat_names <- unique(pi2_for_mdl$malnutrition_category) %>%  sort()
pi2_for_mdl <- map(cat_names, ~ pi2_for_mdl %>% 
                     filter(malnutrition_category == .x) %>% 
                     select(None, Moderate, Severe) %>% 
                     as.matrix())
names(pi2_for_mdl) <- cat_names

pi2_smry <- pi2 %>% 
  group_by(malnutrition_category) %>% 
  summarise_at(vars(pi2_coll12, pi2_coll23), mean) %>% 
  ungroup()

pi2_lng <- pi2 %>% 
  gather("collapsed_categories", "value", pi2_coll12, pi2_coll23) 
plot_dist <- ggplot(pi2_lng, aes(x = value, fill = collapsed_categories)) + geom_histogram() +
  facet_wrap(collapsed_categories ~ malnutrition_category)
plot_dist
```

```{r maln_data}
mort_slct <- mort %>% 
  select(study, malnutrition_category, n = total, r = died, maln = malnutrition_severity,
         col12, col23) %>% 
  inner_join(pi2_smry) %>%
  mutate(pi2 = case_when(
    col12 ~ pi2_coll12,
    col23 ~ pi2_coll23,
    TRUE ~ 1
  ))
  

grp_lbls <-  mort_slct %>% 
  group_by(study, malnutrition_category) %>% 
  mutate(g = seq_along(study)) %>% 
  ungroup() %>% 
  mutate(g_lbl = case_when(
    maln == "None" ~ 1L,
    maln == "Moderate" ~ 2L,
    maln == "Severe" ~ 3L,
    maln == "Non-Severe" ~ 12L,
    maln == "All" ~ 23L)
  )

```


Take the first malnutrition category for each study. Will need to ask group to make a decision on which to use. Then spread the dataframe to wide, so that we have a matrix of N's, events and group labels as per the earlier structure.


```{r labe_groups}
grp_lbls2 <- grp_lbls %>% 
  select(study, malnutrition_category, r, n, g, g_lbl, pi2) %>% 
  distinct(study, g, .keep_all = TRUE)

grp_lbls2n <- grp_lbls2 %>% 
  select(-r, -g_lbl) %>% 
  spread(g, n)

grp_lbls2n <- grp_lbls2 %>% 
  select(-r, -g_lbl) %>% 
  spread(g, n)

grp_lbls2r <- grp_lbls2 %>% 
  select(-n, -g_lbl) %>% 
  spread(g, r)

grp_lbls2g <- grp_lbls2 %>% 
  select(-n, -r) %>% 
  spread(g, g_lbl)

grp_lbls2_col12 <- grp_lbls2 %>% 
  group_by(study) %>% 
  summarise(res = any(12 %in% g_lbl)) %>% 
  pull(res)

```

Check that the restructuring has kept the order of the studies.

```{r check_ordering}

identical(grp_lbls2n %>% select(1:2),
          grp_lbls2g %>%  select(1:2))

identical(grp_lbls2n %>% select(1:2),
          grp_lbls2r %>%  select(1:2))

```


## Try running models on real data

Initially use a fixed proportion 

```{r trymodel}
na <- 3 - apply(grp_lbls2n %>% select(`1`, `2`, `3`) %>% as.matrix(), 1, function(x) x %>% 
        as.integer() %>% 
        is.na() %>% 
        sum())
list_data2 <- list(r = grp_lbls2r %>% select(`1`, `2`, `3`) %>% as.matrix(),
                   n = grp_lbls2n %>% select(`1`, `2`, `3`) %>% as.matrix(),
                   g = grp_lbls2g %>% select(`1`, `2`, `3`) %>% as.matrix(),
                   na = na,
                   pi2 = grp_lbls2$pi2,
                   ng = 3,
                   ns = nrow(grp_lbls2n))

```


```{r runmodel}
mod1 <- jags.model(file = "Supporting/FE_model.txt",
                            data = list_data2, n.chains = 1, n.adapt = 1000)
update(mod1, 1000)

data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod1, c("d"),n.iter = 1000)
summary(LINE.out)
```

## Next step

Next step will be to estimate the proportion of people in category two rather than assume it is fixed. Can use existing data, and sample from Beta distribution. Or can, get input from subject-matter experts for these studies and use that.

In order to prepare for this we developed two models. One model assumes that the proportion of none-moderate which are moderate lies between 0 and 1 and that the proportion of moderate-severe which are moderate lies between 0 and 1 (ie a uniform prior). A second model assumes that the proportion in each category is exchangeable between studies, using a random effects model to estimate the proportion for those studies where it is not recorded.

In these very low MCMC sample models (for speed) they give similar result.

### Model with uniform prior

Add an indicator variables for whether we want proportion in category two of 1 and 2, or of 2 and 3. Where there is no missing data, the proportion of two in 2/3 will be calculated, but this is not used in the code.

In the first instance explore the use of a flat distribution for the proportion in category two

```{r prop_model}
list_data3 <- list_data2
list_data3[["pi2"]] <- NULL
list_data3[["coll12"]] <- grp_lbls2_col12
mod2 <- jags.model(file = "Supporting/FE_model_estimate_prop.txt",
                            data = list_data3, n.chains = 1, n.adapt = 1000)
update(mod2, 1000)

data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod2, c("d", "theta1", "theta2"),n.iter = 1000)
summary(LINE.out)
```

### Model assuming proportion in each category is exchangeable


```{r prop_model_estimate}
pi2_choose <- pi2_for_mdl$`w/a`
pi2_choose12 <- rowSums(pi2_choose[, 1:2])
pi2_choose23 <- rowSums(pi2_choose[, 2:3])

list_data4 <- list_data3
list_data4[["n_complete2a"]]   <- pi2_choose[,2]
list_data4[["n_complete2b"]]   <- pi2_choose[,2]
list_data4[["n_complete12"]] <- pi2_choose12
list_data4[["n_complete23"]] <- pi2_choose23
list_data4[["ns_complete"]] <- length(pi2_choose12)


mod3 <- jags.model(file = "Supporting/FE_model_estimate_prop2.txt",
                            data = list_data4, n.chains = 1, n.adapt = 1000)
update(mod3, 1000)

data(LINE)
LINE$recompile()
LINE.out <- coda.samples(mod3, c("d", "theta1_prop", "theta2_prop"),n.iter = 1000)
summary(LINE.out)
```

# Random effects model

Random effects model did not run. Says indexing is incorrect for w. Will need to go over again.